---
title: "Lsn 16 - MA206Y"
author: "Clark"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.show = 'hide')
```

## Midterm Return and Discussion

## Course Project

## 5.2.2

## Comparing 2 proportions - Theory based test

Null Hypothesis:  The population proportion of boys born to smoking parents is the same as the population proportion of boys born to nonsmoking parents

Alternative Hypothesis: The population proportion of boys born to smoking parents is different from the population proportion of boys born to nonsmoking parents

\begin{align*}
& H_0: \\
& H_a:
\end{align*}

Our parameter of interest is:

\vspace{.5in}

One statistic we could use to estimate this paramter is:

\vspace{.5in}

Recall that we used the applet last class to examine the distribution of this statistic and we said if we squinted our eyes in sort of looked like a Normal Curve:

\vspace{1.in}

If our \textbf{validity conditions} are met, it turns out that we can use a standardized statistic:

\vspace{1.in}

And it has a Normal distribution.  Recall that our standardized statistic gives us a feel for how many standard deviations our observation is from the mean.

From our smoking data we have:

\vspace{1.in}

So our z statistic is:

\vspace{1.in}

We can also find a Confidence Interval in a similar manner.  Note that all the calculations can be done in R using `prop.test()`


```{r}
prop.test(x = c(1975, 255), n = c(3602, 565), alternative = "two.sided", conf.level = 0.95, 
    correct = F)
```

Let's go piece by piece through this code and the output.  Though it might look a bit intimidating, the thing to keep in mind is that the $\chi^2$ value given is $z^2$ found above.

Let's work through 5.3.7