---
title: "Lsn 10 - MA206Y"
author: "Clark"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.show = 'hide')
```

##Admin

##Effect Estimation

Recall that $\pi$ and $\mu$ are called:

\vspace{.5in}

And the goal of statistics is to find something out about these values. The only way we will truly know $\pi$ or $\mu$ is if we sample:

\vspace{.4in}

This is often unrealistic, so we do an experiment and observe $\hat{p}$ or $\bar{X}$.  Do you think that $\hat{p}=\mu$?  Will this ever be true?

\vspace{.5in}



DPE is interested in the percentage of Male cadets that pass the IOCT.  
```{r,warning=FALSE,message=FALSE}
library(tidyverse)
apft<-read.csv("APFT.csv")
males.explore<-apft %>% filter(sex=="M") %>% mutate(pass=IOCT_Time<211)%>%
  group_by(pass)%>%summarise(count=n())
males.explore
```

What is $\hat{p}$ here?  What is $n$?

\vspace{.4in}

Does this give us $\pi$?  What if we wanted to prove that $\pi \neq 0.05$?  What would $H_0$ and $H_a$ be?  

\vspace{.5in}

Using a simulation based method we could do:

```{r}
phat=31/(249+31)
pi=0.05
trials=280
M=1000
outcomes=2
stats=data.frame(trial=seq(1,M),result=NA)
for(i in 1:M){
  simulation=rbinom(trials,outcomes-1,pi)
  stats[i,]$result=sum(simulation)/trials
}

p.value=stats %>% filter(result>phat|result < pi-(phat-pi))%>%summarise(pval=n()/M)
p.value
```

What would our conclusion be if we were willing to risk $\alpha=.05$?


What if we were testing $\pi=0.09$?

\vspace{.5in}

So there's a tipping point where we would switch from rejecting to not rejecting.  Let's go to the applet and see if we can't find that point.  Let's include the $\pi$ we are testing and the proportion of 'as extreme or more extreme' observations:

\vspace{1.5in}

So we're fairly certain that $\pi$ cannot be:

\vspace{.5in}

Note we haven't proven what $\pi$ is, but we've given a range of values that we are fairly confident that our true proportion of Cadet males who fail the IOCT is.

This is a \textbf{Confidence Interval}.  In particular this would be considered a 95\% confidence interval.  What if we were only willing to consider $\alpha=0.01$?  According to our table what would our tipping points be?  

\vspace{.5in}

There's a couple of interpretations of Confidence intervals that are generally accepted to be true.  The first is 'we are XXX percent confident that $\pi$ lies between ZZZ and YYY`.  This is ok.  Another interpretation is if we repeated our experiment over and over and over again, about XXX percent of the time $\pi$ would fall between ZZZ and YYY.  This, in my opinion, is a bit better of an interpretation, but they both are fine for this course.  What is NOT fine is:  XXX percent of our data lies between ZZZ and YYY.  Or, there is a XXX probability that $\pi$ is between ZZZ and YYY.

Just like for hypothesis testing there is a theory based approach for finding confidence intervals.  Note that our validity condition is the exact same as before.  Which are:


\vspace{.5in}

If these are met, then we know:


\vspace{.5in}

For a confidence interval, though, we have to change one small thing.  In particular we aren't testing a hypothesis for a single $\pi_0$ so we need something else in the denominator.  For forming a CI we use:

$$ z=\frac{\hat{p}-\pi}{\sqrt{\hat{p}(1-\hat{p})/n}}$$

This leads to the following formula for a $100(1-\alpha)$ confidence interval:

$$\hat{p}\pm z_{1-\alpha/2}\sqrt{\hat{p}(1-\hat{p})/n}$$ 

Our book refers to $z_{1-\alpha/2}$ as the multiplier.  What I'm saying here is:

\vspace{1.in}

So to find a 95\% CI we would need the spot on our X axis so that the area under the curve is 1-.05/2=.975.  Or

```{r}
qnorm(.975)
```

For a 99\% CI we would need 1-.01/2=.995

```{r}
qnorm(.995)
```

So, four our IOCT data, using $z_{1-\alpha/2}=1.96$ or $\alpha=.05$ find a 95 \% CI 

\vspace{1.in}


We can get the same thing in R by using `prop.test`

```{r}
prop.test(x=31,n=(249+31),alternative="two.sided",conf.level=0.95,correct=F)
```

What would we have to change for a 99\% CI?  What about a 90\% CI?

\vspace{.5in}


The Gallup organizaiton conducted a survey with a random sample of 1,019 adult Americans on Decmeber 10-12, 2010.  They found that 80 \% of the respondents agreed with the statement that the US has a unique character that makes it the greatest country in the world.

In words, what is the population and what is the sample?

\vspace{.5in}

Is $\pi=0.8$ where $pi$ is the proportion of Americans who agree with the statement?

\vspace{.3in}

Is it plausible that $\pi=0.775$?  Use the applet to explore this.  Click on Summary stats.  How many standard deviation is 0.775 from 0.80?

\vspace{.5in}

Note that `qnorm(.975)` yields a value of about 2, so a quick and dirty way to find a 95\% CI is to take $hat{p}\pm 2 sd(\hat{p})$  Using the SD you found in the simulation, do this.

\vspace{.5in}

Interpret this CI

\vspace{.5in}

Can we use a theory based method here?  Why or why not?

\vspace{.3in}

Accourding to the theory based method $sd(\hat{p})=\sqrt{\hat{p}(1-\hat{p})/n}$  What is this value and how does it compare to your simulations?

\vspace{.3in}

Use the theory based method to construct an 85\% CI,  is it wider or narrower than a 95\% CI? 