---
title: "Lsn 17 - MA206Y"
author: "Clark"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Admin

Problem 5.3.7 (a-d, g, h)

\vspace{2.in}

Determine whether hte type of bicycle frame affects travel time to work. Dr. Groves tossed a coin to decide whether he would ride to work on a carbon frame or a steel frame and recorded the weight for 56 days.

Our Null hypothesis and alternative hypothesis are:

\vspace{1.in}

Our response variable is:

\vspace{.5in}

Our parameter of interest is now:

\vspace{.5in}

Note that it would NOT be appropriate to use $\pi_{steel}$ and $\pi_{carbon}$ here.  Why?

\vspace{.5in}

One statistic that we could use to explore our parameter of interest is:

\vspace{.5in}

After we have decided what statistic to use we do our experiment and find the realization of our test statistic, which in this case is:

\vspace{.5in}

Recall that in order to find out how rare our test statistic would be, we need to know the distribution of all possible $\bar{X}_1-\bar{X}_2$ under $H_0$.

```{r}
bike.dat<-read.table("http://www.isi-stats.com/isi/data/chap6/BikeTimes.txt",header=T)
group.means<-bike.dat %>% group_by(Frame)%>%summarize(xbar=mean(Time))
our.stat<-group.means$xbar[1]-group.means$xbar[2] #Sorry!  I don't know a more straight forward way
#to code this!
```

Note that if $H_0$ is true, the labeling doesn't matter. So we can shuffle the labels around:

```{r}
bike.dat.shuff<-bike.dat %>% mutate(shuff.labels=sample(Frame))
```

And under that shuffle we can see what we would get for a statistic
```{r}
group.means<-bike.dat.shuff %>% group_by(shuff.labels)%>%summarize(xbar=mean(Time))
H0.stat<-group.means$xbar[1]-group.means$xbar[2]
```

So we can do this 1000 times to get a feel for the distribution under $H_0$

```{r}
results<-data.frame(repetition=seq(1,1000),stat=NA)
for(j in 1:1000){
  bike.dat.shuff<-bike.dat %>% mutate(shuff.labels=sample(Frame))
  group.means<-bike.dat.shuff %>% group_by(shuff.labels)%>%summarize(xbar=mean(Time))
  H0.stat<-group.means$xbar[1]-group.means$xbar[2]
  results[j,]$stat=H0.stat
}
```

So our simulation distribution is:
```{r}
results %>% ggplot(aes(x=stat))+geom_histogram()+
  geom_vline(xintercept=our.stat,color="red")+
  geom_vline(xintercept=-our.stat,color="red")
```

So, if $H_0$ is true, would it be rare to observe what we actually observed?

\vspace{.5in}

Why is our distribution under $H_0$ centered at 0?

\vspace{.5in}

The same thing can be done with the applet.

Note in either case we can use the 2SD method to construct an approximate 95\% CI.

```{r}
sd(results$stat)
```

Compare to applet


## Practice

Let's work through 6.2.24 on the boards.  