---
title: "Lsn 8 - MA206Y"
author: "Clark"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.show = 'hide')
```

## Admin

Recall that if certain validity conditions were met we could use

$$z=\frac{\hat{p}-\pi_0}{\sqrt{\pi_0(1-\pi_0)/n}}$$

As a test statistic to test whether our true parameter, $\pi=\pi_0$ vs $\pi \neq \pi_0$.  What was the benefit of using this test statistic?

\vspace{.5in}

As it turns out, a similar condition applies when we are looking at the population average, $\mu$.  What is the difference from $\pi$ and $\mu$?

\vspace{.5in}

When the quantitative variable is symmetric or we have at least 20 observations and the sample distribution is not strongly skewed we can use:

$$ t=\frac{\bar{x}-\mu_0}{s/\sqrt{n}}$$

As our test statistic.  And in this case we know the distribution of $t\sim t_{n-1}$.  Below we can see the difference from a t distribution and a normal distribution.  Play around with $n$.  What happens as $n$ gets really big?

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
n=5
sim.data.t=data.frame(sim.tstat=rt(1000,n-1))
sim.data.z=data.frame(sim.zstat=rnorm(1000))
ggplot(sim.data.t, aes(x=sim.tstat)) +
  stat_function(fun = dt, args = list(df=n-1), colour = "blue",lwd=2)+
  stat_function(fun=dnorm,colour="red",lwd=2)
```

So, if we want to test hypothesis like $H_0: \mu=\mu_0$ vs $H_a: \mu\neq \mu_0$ we could form a t-statistic and compare it to a t-distribution.  (Note there's a simulation way to do this, but I don't think it's overly insightful in this case).


I want to demonstrate to the BTO that Cadets are getting aren't getting enough sleep.  In particular I want to show that the average amount of sleep that Cadets are not getting 6 hours of sleep.

Based on this question, what is the population of interest?  What is the parameter?  What symbol do we use as a parameter?

\vspace{.5in}

Express the null and alternative hypothesis for investigation this question

\vspace{.5in}

Ideally, how would we sample Cadets to investigate this question?

\vspace{.5in}

I decide to just sample Cadets in my class, what kind of sample is this?  Is this potentially a biased study?

\vspace{.5in}

How much sleep (to the nearest quarter hour) did you get last night?  Combine your answer with those of your classmates.

Load the data into R and create a histogram of your data.  Describe the distribution.  What is the center of the distribution?  What do we call this?

\vspace{.5in}

How many students are above the sample mean, how many are below?

\vspace{.5in}

What is the \textbf{sample standard deviation} for our study?  What letter do we use for this?

\vspace{.5in}

Are our validity conditions met to use a theory based approach?

\vspace{.5in}

Let's assume that they are.  Form our \textbf{t statistic}.  What is the distribution of our t-statistic asusming our null hypothesis is true?

\vspace{.5in}

Draw a t distribution.  Where does OUR value lie?  If the null hypothesis was true would our value be rare?

\vspace{.5in}

Just like `pnorm`  R has a command called `pt` that integrates for us (HOORAY).  Using this, find the p-value for our study.

\vspace{.5in}

Let's do the same thing with a built in r function called `t.test`.  Run `?t.test` in your console.  What inputs do we need to put in?
\vspace{.5in}

If we conclude that Cadets are not getting 6 hours of sleep, what is the probability we have committed an error?

\vspace{.5in}

In statistics there are two types of errors we need to be concerned with, Type I and Type II.  Oftentimes Cadets find it useful to think of Type I and Type II errors as Jurys.  Let's look at Table 2.9 and Table 2.10

\vspace{1.5in}

Oftentimes we want to FIX our Type I error and use this create a decision rule.  Note that this is done BEFORE we take our sample.  For instance, if we set $\alpha=0.1$ we are saying:

\vspace{1.in}

The notion of Type II error is a bit more complicated.  But perhaps a drawing will help:

\vspace{1.in}

If we are testing:
\begin{align*}
&H_0: \mu = 6 \\
&H_a: \mu < 6
\end{align*}
And we set $\alpha=0.1$ what values of our t-statistic would we reject $H_0$ for?

\vspace{1.in}


What about
\begin{align*}
&H_0: \mu = 6 \\
&H_a: \mu > 6
\end{align*}
And we set $\alpha=0.1$ what values of our t-statistic would we reject $H_0$ for?

\vspace{1.in}

If we are testing:

\begin{align*}
&H_0: \mu = 6 \\
&H_a: \mu < 6
\end{align*}

And we set $\alpha=0.2$ what values of our t-statistic would we reject $H_0$ for?

\vspace{1.in}


What about
\begin{align*}
&H_0: \mu = 6 \\
&H_a: \mu \neq 6
\end{align*}

And we set $\alpha=0.2$ what values of our t-statistic would we reject $H_0$ for?


\vspace{1.in}

Next lesson will will discuss parameter estimation.  Recall that there is a difference between $\bar{X}$ and $\mu$.  Why can't we say $\bar{X}=\mu$?  

\vspace{.5in}

While we don't \textit{know} $\mu$ from our experiment, perhaps $\bar{X}$ can tell us plausible values of $\mu$.  From our study today is it likely that $\mu=8$?  What about $\mu=5$?